### 2018-06-03 index.rmd
adding hesscl.com link to navbar of two pages

---
title: "Estimating Neighborhood Rents with Scraped Data"
output: 
  html_document:
      includes:
        in_header: header.html
        after_body: footer.html
---

<br>

This project is under construction and subject to change at any moment!

<br>

### Contents

[Working paper on estimating Seattle neighborhood rent indicators](./manuscript/index.html)

[`flexdashboard` for working paper's model diagnostics and forecasts](./dashboard/index.html)

[GitHub Repository](https://github.com/hesscl/smooth-sea/)

<br>

### See also

[Helena](http://helena-lang.org) - Web automation for end users

<br>

### Contact

Email Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.