### 2018-06-03 index.rmd
update to main index, also settled navbar for now

---
title: "Estimating Neighborhood Rents with Scraped Data"
output: 
  html_document:
      includes:
        in_header: header.html
        after_body: footer.html
---

<br>

This project is under construction and subject to change at any moment!

<br>

### Objective

The goal of this research is to understand the spatial and temporal dynamics of rent estimates in Seattle using scraped rental housing listings from Craigslist. I use a series of data visualizations to describe the overall datam, after which I model a panel of quarterly census tract estimates to explain the dynamics observed so far and forecast a surface of rent values for the current quarter.

### Contents

[Working paper on estimating Seattle neighborhood rent indicators](./manuscript/index.html)

[`flexdashboard`](./dashboard/index.html) for working paper's model diagnostics and forecasts

[GitHub Repository](https://github.com/hesscl/smooth-sea/) with source code for manuscript and results dashboard

<br>

### See also

[Helena](http://helena-lang.org) - Web automation for end users

<br>

### Contact

Email Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.