### 2018-06-04 manuscript/index.Rmd

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess"
date: "June 2, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: "header.html"
      after_body: "../footer.html"
abstract: "Online classified ads are a key source of information about local rental housing markets. This research asks what temporal and spatial dynamics explain neighborhood variations in one-bedroom rents in Seattle, WA, and contributes a novel approach to forecasting quarterly neighborhood rent levels using Bayesian models estimated with integrated nested Laplace approximations (INLA). All models were trained on data for 2017 Q2 through 2018 Q1, with forecasts for 2018 Q2 compared against the observed-to-date data for the current quarter (i.e. 2018 Q2). The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $90 for the training data and $145 for the test data."
---

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #4b2e83;
}
</style>

```{r setup, include=FALSE}
#packages
library(tidyverse) #need dev version of ggplot
library(forcats)
library(haven)
library(lubridate)
library(haven)
library(zoo)
library(ggthemes)
library(RColorBrewer)
library(sp)
library(spdplyr)
library(geosphere)
library(rgdal)
library(rgeos)
library(sqldf)
library(spdep)
library(viridis)
library(latex2exp)

#load ACS data extract (codebook in folder)
census <- read_csv(file = "../input/acsExtract.csv")
censusRent <- read_csv(file = "../input/acsRent.csv")
census <- inner_join(census, censusRent) #1:1

#read in tract shapefiles for seattle
sea_shp <- readOGR(dsn = "../input/sea_tract_2010/sea_tract_2010.shp",
                  layer = "sea_tract_2010",
                  GDAL1_integer64_policy = TRUE,
                  stringsAsFactors = F)

#go ahead and save separate versions for graphics later
tract2000 <- readOGR(dsn = "../input/sea_tract_2000/sea_tract_2000.shp",
                  layer = "sea_tract_2000",
                  GDAL1_integer64_policy = TRUE,
                  stringsAsFactors = F) 
tract2010 <- sea_shp

#source data_load()
source("../helpers/load_data.R")

#use sourced function to load listing summaries
tract <- load_data()
```

<br>
<br>

## Introduction

Demand for rental housing units in Seattle, WA surged in the past decade, due to relative scarcity in available housing amidst a period of historic population growth. Growing shares of households are unable to buy a home short of "driving till they qualify", and among households who rent a substantial share pay 30%, if not more of their incomes on their housing costs. The rising cost of housing therefore has a direct role in neighborhood processes of "demographic inversion" like urban gentrification and suburbanizing poverty. American Community Survey data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more. This publication lag for ACS data prohibits analyzing any market dynamics beyond long-term trends, particularly for estimates of smaller areas of aggregation like neighborhoods where ACS uses 5-year rolling-average methods as well. While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties. This would-be problem for academic researchers creates an opportunity for applying web scraping methods to generate a novel data on housing availability and costs.

This study aims to demonstrate why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography through a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA. Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data. In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub. I start by briefly reviewing some background relevant to studying neighborhood rental housing. I then outline a process for automated web scraping of a rental listing source like Craigslist using a program called Helena, and compare aggregates from the collected rental listings with other data to assess rent agreement in measures. After describing the data and process, I present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data. I conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.

<br>

## Background

Existing public data sources like the American Community Survey or decennial census theoretically have two sources of error relative to the ``true'' median rent in a given overall market and to an even greater degree among neighborhood approximations like census tracts. Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication (i.e. ACS estimates are usually at least 2 years old, decennial many more). This feature of Census data is salient given newer housing units' generally higher rents and greater-than-average vacancy, both factors which lead to underestimating the present levels of rents for available housing units when either the ACS or decennial census are used as a current indicator. For already-leased rental housing units that are sampled, the temporal lag generates measurement error such that data on a given building's current asking rents may be underestimated relative to current tenants' payment levels.

For-profit data sources play a role in filling the information gap about rental housing costs, though they tend to possess their own unique shortcomings. These data sources tend to rely on survey methodologies for data collection, and ask landlords in a given market to outline the average cost for units of a given bedroom size or other specification. In exchange for providing detailed information for their database, the rental data provider may give participating landlords and property management companies access to a slice of information generated via accumulated data. These data tend to be rather opaque in their collection methodologies, with larger or corporate-owner properties most likely to appear in these data relative to independently-owned (i.e. mom and pop) rental units.

<br>

## Data

### Web scraping using Helena

[Helena](http://helena-lang.org) is a high-level programming language for end users to automate web scraping as a method of data collection. Users write programs in an extension for Google Chrome by demonstrating the first row of data they want to collect using clicks, form interactions and highlighting values. The extension translates an initial demonstration into a starting script written in Helena, which a user can further modify or set to run on a regular schedule. Though Helena programs collect data roughly at the speed they are recorded at, users can run multiple programs in parallel to increase the volume of data collected. Data are stored on an AWS server where users can later query either the full table of scraped data or a subset scraped within a specified number of hours.

Helena learns where features of are through the user interface (UI) rather than the underlying document object model (DOM). Though the HTML code of a site may regularly change over the course of regular data scraping, websites less frequently undergo drastic overhauls to what their users interact with. Helena also uses skip blocks to avoid re-scraping the an already-collected item. These two aspects of Helena provide its programs robustness to common headaches encountered with web scraping by default. Overall, Helena's programming-by-demonstration method reduces time needed building a scraper from scratch, giving social scientists the capacity to quickly scrape a single table of data or build ongoing data processes.

For the UW team's data collection, a Helena program runs 24-hours a day and scrapes listings from the Seattle Craigslist. It is important to note that web scraping does violate Craigslist's Terms of Service, and that we are using all scraped data exclusively for academic research. For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf).

<br>

### Data Processing

The key steps of any data process can be distilled into three tasks: a.) _extract_ a table of interest from a specified origin source, b.) _transform_ it to support the desire analyses and c.) _load_ the transformed data to a storage destination for future use. In the present case, a program extracts scraped listings observed since the database 
<br>

### Data Description

```{r map_point_density, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../../data/cl/craigslistDB.sqlite")){
  source("../helpers/map_point_density.R") #if config, refresh plot pdf
}
```

![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)

```{r rent_lineplot, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../../data/cl/craigslistDB.sqlite")){
  source("../helpers/rent_lineplot.R") #if config, refresh plot pdf
}
```

![Figure 2. Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018. Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual and ACS are annual (i.e. respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)

```{r rent_map_across, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../../data/cl/craigslistDB.sqlite")){
  source("../helpers/rent_map_across.R") #if config, refresh plot pdf
}
```

![Figure 3. Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](../output/rent_map_across_1B.png)

```{r rent_scatter_across, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../../data/cl/craigslistDB.sqlite")){
  source("../helpers/rent_scatter_across.R") #if config, refresh plot pdf
}
```

![Figure 4. Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](../output/rent_scatter_across_1B.png)

<br>

## Models

```{r construct panel (some missingness), echo=FALSE, message=FALSE, warning=FALSE}
#ensure order of census df matches shapefile
census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]

#clean a few fields we might use
census <- census %>%
  mutate(nHU = AF7PE001,
         idtract = row_number(),
         pOwnoccHU = AF7PE002/AF7PE001,
         medHUVal = AF9LE001)

#store NULL object to bind iterated dfs into
panel <- NULL

#for each unique quarter in the CL data
for(i in unique(as.character(tract$listingQtr))){
  period <- tract %>% #start pipe with tract, end with period
    mutate(listingQtr = factor(listingQtr, ordered = T)) %>% #make factor
    filter(listingQtr == i) %>% #filter to qtr i
    right_join(census) %>% #right join to census, allows missingness
    ungroup() %>% #remove grouping by tract
    mutate(listingQtr = i, #listingQtr == current iteration
           Qtr = gsub(pattern = "^\\d{4} ", replacement = "", x = listingQtr)) #subset str to Q1,2,3,4
  panel <- bind_rows(panel, period) #append to panel object
}

#establish object for modeling based off of adjusting `panel`
sea_df <- panel %>% 
  mutate(actualRent = med1B, #first preserve the observed / observed-to-date
         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>% #set to NA for INLA to forecast current qtr
  mutate(p1B = n1B/nListings) %>% #compute share of 1B listings
  arrange(listingQtr, GISJOIN) #arange qtr, tract
```

```{r run models, echo=FALSE, message=FALSE, warning=FALSE}
library(INLA)

#create adjacency matrix from shapefile
sea_adj <- poly2nb(sea_shp)

#create neighbor file that INLA takes for hyperparameter args
nb2INLA("../output/seatract.graph", sea_adj)
sea_df$idqtr <- factor(sea_df$listingQtr, ordered = T)
sea_df$idqtr1 <- sea_df$idqtr
sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$Qtr)
sea_df$idtract1 <- sea_df$idtract
sea_df$Qtr1 <- sea_df$Qtr

#fixed effect for qtr only
form.int <- log(med1B) ~ 1 + Qtr

m.int <- inla(form.int, 
             family = "normal", 
             data = sea_df,
             control.predictor = list(compute = TRUE),
             control.compute = list(dic = TRUE, waic = TRUE))

summary(m.int)
sea_df$int_Med <- m.int$summary.linear.predictor[, "0.5quant"]
sea_df$int_SD <- m.int$summary.linear.predictor[, "sd"]
sea_df$int_postWidth <- m.int$summary.linear.predictor[, "0.975quant"] - m.int$summary.linear.predictor[,"0.025quant"]

#fixed qtr + iid tract random effect
form.ns <- log(med1B) ~ 1 + Qtr +
   f(idtract, model = "iid")

m.ns <- inla(form.ns, 
              family = "normal", 
              data = sea_df,
              control.predictor = list(compute = TRUE),
              control.compute = list(dic = TRUE, waic = TRUE))

summary(m.ns)
sea_df$ns_Med <- m.ns$summary.linear.predictor[, "0.5quant"]
sea_df$ns_SD <- m.ns$summary.linear.predictor[, "sd"]
sea_df$ns_postWidth <- m.ns$summary.linear.predictor[, "0.975quant"] - m.ns$summary.linear.predictor[,"0.025quant"]
sea_df$ns_Eff <- m.ns$summary.random$idtract[sea_df$idtract, "0.5quant"]

#IID tract random effect + AR(1) process
form.nsar1 <- log(med1B) ~ 1 +
  f(idtract, model = "iid") +
  f(idqtr, model = "ar1") +
  f(idqtr1, model = "iid")

m.nsar1 <- inla(form.nsar1, 
               family = "normal", 
               data = sea_df,
               control.predictor = list(compute = TRUE),
               control.compute = list(dic = TRUE, waic = TRUE))

summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
sea_df$nsar1_Med <- m.nsar1$summary.linear.predictor[, "0.5quant"]
sea_df$nsar1_SD <- m.nsar1$summary.linear.predictor[, "sd"]
sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
sea_df$nsar1_Eff <- m.nsar1$summary.random$idtract[sea_df$idtract, "0.5quant"]

#spatial random effect model (BYM 1991) with AR(1)
form.bym <- log(med1B) ~ 1 +
  f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
    scale.model = T,
    graph = "../output/seatract.graph") +
  f(idqtr, model = "ar1") +
  f(idqtr1, model = "iid")

m.bym <- inla(form.bym, 
             family = "normal", 
             data = sea_df,
             control.predictor = list(compute = TRUE),
             control.compute = list(dic = TRUE, waic = TRUE))

summary(m.bym)
sea_df$bym_Med <- m.bym$summary.linear.predictor[, "0.5quant"]
sea_df$bym_SD <- m.bym$summary.linear.predictor[, "sd"]
sea_df$bym_postWidth <- m.bym$summary.linear.predictor[, "0.975quant"] - m.bym$summary.linear.predictor[,"0.025quant"]
sea_df$bym_Eff <- m.bym$summary.random$idtract[sea_df$idtract, "0.5quant"]

#spatio-temporal smoothing model, with AR(1) and linear space-time interaction (Bernadelli 1995)
form.spt <- log(med1B) ~ 1 + 
  f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
    scale.model = T,
    graph = "../output/seatract.graph") +
  f(idqtr, model = "ar1") +
  f(idqtr1, model = "iid") +
  f(idtractqtr, model = "iid")

m.spt <- inla(form.spt, 
              family = "normal", 
              data = sea_df,
              control.predictor = list(compute = TRUE),
              control.compute = list(dic = TRUE, waic = TRUE),
              control.inla = list(correct = TRUE, correct.factor = 10))

summary(m.spt)
sea_df$spt_Med <- m.spt$summary.linear.predictor[, "0.5quant"]
sea_df$spt_SD <- m.spt$summary.linear.predictor[, "sd"]
sea_df$spt_postWidth <- m.spt$summary.linear.predictor[, "0.975quant"] - m.spt$summary.linear.predictor[,"0.025quant"]
sea_df$spt_Eff <- m.spt$summary.random$idtract[sea_df$idtract, "0.5quant"]

#overfitting? too much bumpiness qtr-to-qtr

#### Maps of Model Output ------------------------------------------------------------------

sea_df <- sea_df %>%
  mutate_at(.vars = vars(matches("_Med")),
            .funs = exp) %>%
  mutate_at(.vars = vars(matches("_SD")),
            .funs = exp) %>%
  group_by(GISJOIN) %>% 
  arrange(GISJOIN, listingQtr) %>%
  mutate(spt_traj = spt_Med - lag(spt_Med, 4)) %>%
  ungroup %>%
  group_by(GISJOIN) %>%
  mutate(spt_traj = sum(spt_traj, na.rm=T)) %>%
  ungroup() %>%
  mutate(bym_cuts = cut(bym_Med, 6),
         spt_cuts = cut(spt_Med, 6))

sea_shp@data$id <- rownames(sea_shp@data)
sea_shp@data <- left_join(sea_shp@data, sea_df)
sea_f <- fortify(sea_shp)
sea_f <- inner_join(sea_f, sea_shp@data,"id")

rmse <- function(error){sqrt(mean(error^2))}
mae <- function(error){mean(abs(error))}

#table of fit statistics
fit_stats <- sea_df %>%
  mutate(train_test = ifelse(listingQtr == as.yearqtr(Sys.Date()), "Test", "Training")) %>%
  mutate(int_err = actualRent - int_Med,
         ns_err = actualRent - ns_Med,
         nsar1_err = actualRent - nsar1_Med,
         bym_err = actualRent - bym_Med,
         spt_err = actualRent - spt_Med) %>% 
  group_by(train_test) %>%
  summarize(int_rmse = rmse(int_err[!is.na(int_err)]),
            int_mae = mae(int_err[!is.na(int_err)]),
            ns_rmse = rmse(ns_err[!is.na(ns_err)]),
            ns_mae = mae(ns_err[!is.na(ns_err)]),
            nsar1_rmse = rmse(nsar1_err[!is.na(nsar1_err)]),
            nsar1_mae = mae(nsar1_err[!is.na(nsar1_err)]),
            bym_rmse = rmse(bym_err[!is.na(bym_err)]),
            bym_mae = mae(bym_err[!is.na(bym_err)]),
            spt_rmse = rmse(spt_err[!is.na(spt_err)]),
            spt_mae = mae(spt_err[!is.na(spt_err)])) %>%
  dplyr::select(train_test, ends_with("_rmse"), ends_with("_mae"))
```

`r knitr::kable(fit_stats)`

<br>

## Results

<br>

## Discussion

<br>

## References